Comparison of FDR methods on GWAS with SummarizedBenchmark

### 2017-10-29 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Add parallel buildBench, new plotting functions

# comparison tools/functions
library(SummarizedBenchmark)
sourceDirectory("benchmark-fdr/datasets/R/")
```

@@

### experimenting on a subset of the full dataset for now
bmi <- bmi[1:500000,]
```

@@

```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
strat_hist(bmi, pval="pval", covariate="ind_covariate", maxy=3)
rank_scatter(bmi, pval="pval", covariate="ind_covariate")
```

@@

```{r, GWAS1-sb, results="hide"}
resfile <- paste0(resdir, "bmi_summarizedBenchmark_",
                  nrow(bmi), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  sb <- bd %>% buildBench(data=bmi)
  metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
  save(sb, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(bmi), " SNPs."))
}else{
  load(resfile)
}

@@

# estimate perf metrics and store in tidy format for plotting
pf <- estimatePerformanceMetrics(sb, alpha=seq(0.01, 0.10, 0.01), tidy=TRUE)
```

@@

```{r, GWAS1-plot, results="hide", width=15, height=15}
# plot adjusted p-values by effect_size (diagnostic from ASH paper)
plotsim_single(data.frame(assay(sb)), bmi, strat = "effect_size")

# plot adjusted p-values by independent covariate
plotsim_single(data.frame(assay(sb)), bmi, strat = "ind_covariate")

# alpha versus nrejects
plotsim_average(pf, met="rejections", merge_ihw = TRUE) +
  ggtitle("Number of rejections for BMI dataset")

# alpha versus nrejects (excluding "unadjusted")
plotsim_average(pf, met="rejections", filter_set = "unadjusted", merge_ihw = TRUE) +
  ggtitle("Number of rejections for BMI dataset")

# upset plot
plotMethodsOverlap(sb, alpha=0.1, nsets=ncol(sb))

# upset plot excluding "unadjusted"
plotMethodsOverlap(sb[,-1], alpha=0.1, nsets=ncol(sb)-1)
```

>>>>

# comparison tools/functions
library(SummarizedBenchmark)
sourceDirectory("benchmark-fdr/datasets/R/")

# set up parallel backend
library(BiocParallel)
cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
multicoreParam <- MulticoreParam(workers = cores)
```

@@

### experimenting on a subset of the full dataset for now
bmi <- bmi[1:750000,]
```

@@

```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
strat_hist(bmi, pvalue="pval", covariate="ind_covariate", maxy=3)
rank_scatter(bmi, pvalue="pval", covariate="ind_covariate")
```

@@

```{r, GWAS1-sb, results="hide"}
resfile <- paste0(resdir, "bmi_summarizedBenchmark_",
                  nrow(bmi), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  sb <- bd %>% buildBench(data=bmi, parallel=TRUE, BPPARAM=multicoreParam)
  metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
  save(sb, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(bmi), " SNPs."))
}else{
  load(resfile)
}

@@

# estimate perf metrics 
sb <- estimatePerformanceMetrics(sb, alpha=seq(0.01, 0.10, 0.01), addColData=TRUE)
```

@@

```{r, GWAS1-plot, results="hide", width=15, height=15}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sb)

# Add ind covariate to rowData
rowData(sb)$ind_covariate <- bmi$ind_covariate

rejection_scatter_bins(sb, covariate="ind_covariate", bins=4)

# upset plot (only plot IHW for the plotted alpha)
rmv.ihw <- which(grepl("ihw", colnames(sb)) &
                              colData(sb)$param.alpha != 0.05)

plotMethodsOverlap(sb[,-rmv.ihw], 
                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw),
                   order.by="freq", decreasing=TRUE)

# upset plot excluding "unadjusted"
plotMethodsOverlap(sb[,-c(1,rmv.ihw)], 
                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw)-1,
                   order.by="freq", decreasing=TRUE)
```
